---
output:
  html_document: default
  pdf_document: default
---


# Analysis code

#### 1. Load in packages and data

The "pwr" package needs to be installed to the PC's local in order to run this script. 

```{r library_data}

# install package - pwr

library(pwr)
library(broom)
library(dplyr)
library(stats)
library(ggplot2)
library(tidyverse)

# you will need to add extra packages in here to do the rest of your analyses

demo <- read_csv("demographics_2021_pilot.csv")
mslq <- read_csv("MSLQ_2021_pilot.csv")

```

#### 2. Clean up the data


```{r data_clean}

demo_final <- demo %>% 
  group_by(user_id, q_id) %>% 
  filter(session_id == min(session_id), endtime == min(endtime)) %>% 
  filter(row_number() == 1) %>% 
  ungroup() %>% 
  filter(user_status %in% c("guest", "registered")) %>%
  select(user_id, user_sex, user_age, q_name, dv) %>%
  pivot_wider(names_from = q_name, values_from = dv)

mslq_final <- mslq %>% 
  group_by(user_id, q_id) %>% 
  filter(session_id == min(session_id), endtime == min(endtime)) %>% 
  filter(row_number() == 1) %>% 
  ungroup() %>% 
  filter(user_status %in% c("guest", "registered")) %>%
  select(user_id, user_sex, user_age, q_name, dv) %>%
  arrange(q_name) %>%
  pivot_wider(names_from = q_name, values_from = dv)

```

#### 3. Join together the data files by their common columns

```{r join_data}

# merge data frames
joined_data <- as_tibble(merge(demo_final, 
                               mslq_final))

```


#### 4. Use select to retain only the variables you need for your chosen research design (including the user ID).

```{r select_cols}

# select the gender column, and MSLQ variables
selected_data <- select(joined_data,
                        user_id, user_sex,
                        intrinsic_1:intrinsic_4,
                        help_1:help_4)

```

#### 5. If necessary, use a filter to retain only the observations you need. For example, you might need to delete participants above a certain age, or only use mature students etc.

As discussed in Question #3, observations with an empty field for user_sex is dropped.

```{r filter_option}

# convert to data.frame and drop NaNs
data <- data.frame(drop_na(selected_data))

```


#### 6. Use `summary()` or `str()` to check what type of variable each variable is. Recode any necessary variables as factors and, if you would like to, change numeric codes (e.g., 1 for native speaker) into words to make it easier to read the output. 

```{r check_type}

# data info
summary_stats <- summary(data)

# MSLQ variables
subscales <- c("intrinsic", 
               "help")
help      <- c(names(data)[str_detect(names(data), 
                                      "help")])
intrinsic <- c(names(data)[str_detect(names(data), 
                                      "intrinsic")])

# group non-female observations
data <- mutate(data,
               user_sex = case_when(user_sex %in% c("nonbinary",
                                                    "na") ~ "non-female",
                                  user_sex == "male"      ~ "non-female",
                                  user_sex == "female"    ~ "female"))

# group by count
data_count <- (data %>% 
                 count(user_sex))

# restructure
data_distribution <- pivot_longer(data,
                                  cols=c(intrinsic, help),
                                  names_to="item",
                                  values_to="response")

# gender variables
gender   <- c(unique(data_distribution$user_sex))
n_gender <- c(sum(data_count$n[data_count$user_sex == "female"]),
              sum(data_count$n[data_count$user_sex != "female"]))

# plotting variables
black    = "#000000"
grey     = "#999999"
yellow   = "#E69F00"

# plot gender counts
# jpeg(file="plots/count.jpeg")
barplot(data_count$n, main="Figure 1: Sample size", 
        names.arg=data_count$user_sex,
        ylab="n", xlab="user_sex",
        col=c(grey, yellow, yellow, yellow))
legend("topright", 
       c(sprintf("%s = %i", gender[1], n_gender[1]), 
         sprintf("%s = %i", gender[2], n_gender[2])), 
       fill = c(grey, yellow))
# dev.off()

# plot data distribution
# jpeg(file="plots/distribution.jpeg")
ggplot(data=data_distribution,
       mapping=aes(x=response, color=user_sex, fill=user_sex)) + 
  geom_histogram(binwidth=.5, position="dodge") + 
  facet_wrap(facet=vars(item)) +
  scale_color_manual(values=c(grey, yellow)) +
  scale_fill_manual(values=c(grey, yellow)) + 
  theme(legend.position="top") +
  ggtitle("Figure 2: Data distribution")
# dev.off()

```

#### 7. Calculate the mean score for each participant for each sub-scale. There are a few ways you can do this but helpfully the Experimentum documentation provides example code to make this easier - shown below. For now, you just need to adapt the below code for the variables you need. You may also want to change the `na.rm = TRUE` for the calculation of means depending on whether you want to only include participants who completed all questions.

At the top of the code chunk below, change `eval = FALSE` to `eval = TRUE` once you have amended your code. The reason it is currently set to FALSE is to allow the file to knit. `eval = FALSE` says ignore the code chunk. `eval = TRUE` says run the code chunk

```{r desc, eval = TRUE}

dat_means <- data %>% # change data to the name of the data object you want to work from
  gather(var, val, intrinsic_1:intrinsic_4) %>% # change question_1:question_5 to select the questions for your 1st sub-scale 
  group_by_at(vars(-val, -var)) %>% # group by everything except the val and var columns, don't change this 
  summarise(intrinsic_mean = mean(val, na.rm = TRUE)) %>% # change anxiety_mean to the name of your 1st sub-scale
  ungroup() %>% # always ungroup! 
  gather(var, val, help_1:help_4) %>% # change question_1:question_5 to select the questions for your 2nd scale
  group_by_at(vars(-val, -var)) %>% 
  summarise(help_mean = mean(val, na.rm = TRUE)) %>% # does not return sums with missing items 
  ungroup()

```

#### 8. You now have the dataset in the format that you need for analysis. You could have actually combined all of the above steps together in one mega pipe-line of code if you felt confident but when starting it helps to break it down. Next, you should visualise the data for each analysis.

T-test visualisation: Based on the graphical representation of our data, the outliers illustrated in Figure 3 were removed before conducting further analysis.

```{r ttest_vis}

# generate dataframe to restructure help_mean and intrinsic_mean into the same column
dat_means_grouped <-  pivot_longer(dat_means,
                      cols=c(intrinsic_mean, help_mean),
                      names_to="subscale",
                      values_to="mean")

# jpeg(file="plots/boxplots.jpeg")
ggplot(data=dat_means_grouped,
       mapping=aes(x=subscale, y=mean, color=user_sex, fill=user_sex)) + 
  geom_violin(trim=FALSE, alpha=0.5) +
  geom_boxplot(width=0.07, position=position_dodge(width=0.9)) +
  scale_color_manual(values=c(black, black)) +
  scale_fill_manual(values=c(grey, yellow)) + 
  theme(legend.position="right") +
  ggtitle("Figure 3: Violin and boxplots separately for gender and MSLQ subscales") 
# dev.off()

# remove outliers
female_outliers      <- boxplot(subset(subset(dat_means_grouped, user_sex == "female"), subscale == "help_mean")$mean, plot=FALSE)$out
non.female_outliers  <- boxplot(subset(subset(dat_means_grouped, user_sex != "female"), subscale == "intrinsic_mean")$mean, plot=FALSE)$out

dat_means_grouped <- dat_means_grouped[-which(subset(subset(dat_means_grouped, user_sex == "female"), subscale == "help_mean")$mean %in% female_outliers),]
dat_means_grouped <- dat_means_grouped[-which(subset(subset(dat_means_grouped, user_sex != "female"), subscale == "intrinsic_mean")$mean %in% non.female_outliers),]

dat_means <- dat_means[-which(
  subset(dat_means, user_sex == "female")$help_mean
  %in% 
    boxplot(subset(dat_means, user_sex == "female")$help_mean, plot=FALSE)$out),]

dat_means <- dat_means[-which(
  subset(dat_means, user_sex != "female")$intrinsic_mean
  %in% 
    boxplot(subset(dat_means, user_sex != "female")$intrinsic_mean, plot=FALSE)$out),]

```

Correlation visualisation

```{r corr_vis}

female_data     <- subset(dat_means, 
                          user_sex == "female")
non.female_data <- subset(dat_means,
                          user_sex == "non-female")

# jpeg(file="plots/correlation.jpeg")
ggplot(data=female_data, 
       aes(x=intrinsic_mean, 
           y=help_mean)) + 
  geom_point() +
  ggtitle("Figure 4: Scatterplot between help-seeking and intrinsic goal orientation responses in females")
# dev.off()

```

#### 9. Now you should check that the data meets the assumptions of the tests you want to conduct.

T-test assumptions: Normality

```{r ttest_assump}

# shapiro-wilk test for normality
# separated for gender and subscale

gender_      <- vector(mode="list")
subscale_    <- vector(mode="list")
p.value_     <- vector(mode="list")
d.statistic_ <- vector(mode="list")

count = 1

for (i in seq_along(gender))
  {
  for (j in seq_along(subscales))
    {
    
    vec <- 
      subset(dat_means_grouped,
                user_sex == gender[i]) %>%
      filter(., 
             grepl(subscales[j], 
                   subscale))
    
    vec_sw <- shapiro.test(vec$mean)
    
    gender_[count]      <- gender[i]
    subscale_[count]    <- subscales[j]
    d.statistic_[count] <- vec_sw$statistic
    p.value_[count]     <- vec_sw$p.value
    
    count = count + 1
    
    }
  }

sw_results     <- tibble(gender_, 
                         subscale_,
                         d.statistic_, 
                         p.value_)

gender_results <- tibble("gender"            = gender,
                         "num. observations" = n_gender)

# jpeg(file="plots/mean_distribution.jpeg")
ggplot(data=dat_means_grouped,
       mapping=aes(x=mean, color=user_sex, fill=user_sex)) +
  geom_histogram(binwidth=1, position="dodge") +
  facet_wrap(facet=vars(subscale)) +
  scale_color_manual(values=c(grey, yellow)) +
  scale_fill_manual(values=c(grey, yellow)) +
  theme(legend.position="right") +
  ggtitle("Figure 5: Mean data distribution")
# dev.off()
 
```

Correlation assumptions: Normality, linearity and homoscedasticity

```{r cor_assump}

# test for linearity
lingress_model <- lm(data=female_data,
                     help_mean ~ intrinsic_mean)

# jpeg(file="plots/data_linearity.jpeg")
plot(lingress_model, 1)
title(main = "Figure 6: Visualisation for linearity",
      cex.main = 1, line = 1.25, adj = 0)
# dev.off()

# test for homoscedasticity
# jpeg(file="plots/data_homoscedasticity.jpeg")
plot(lingress_model, 3)
title(main = "Figure 7: Visualisation for homoscedasticity",
      cex.main = 1, line = 1.25, adj = 0)
# dev.off()

```

``` {r m.clean_vars, eval=TRUE}

# remove raw data objects
rm(demo, 
   mslq,
   demo_final,
   mslq_final,
   joined_data,
   selected_data)

rm(vec,
   vec_sw,
   gender_,
   p.value_,
   subscale_,
   d.statistic_,
   data_count,
   data_distribution,
   lingress_model)

rm(i, j,
   count,
   gender,
   n_gender,
   help, 
   intrinsic)

```

#### 10. Finally, you can conduct your statistical analyses. Don't forget to calculate effect sizes for the t-tests!

Between-conditions Welch T-test.

```{r ttest_analysis}

# cohen's d
mean_diff = mean(female_data$intrinsic_mean) - mean(non.female_data$intrinsic_mean)
sum_var   = var(female_data$intrinsic_mean) + var(non.female_data$intrinsic_mean)
d <- mean_diff/sqrt((sum_var)/2)

# power analysis
pwr <- power.t.test(n           = NULL,
                    delta       = mean_diff,
                    # sd          = 1,
                    sig.level   = 0.05,
                    power       = 0.80,
                    type        = "two.sample",
                    alternative = "one.sided")

effect_sizes = c(seq(0.1,1.5,0.1))
ptab        <- cbind(NULL, NULL) 

for (i in effect_sizes) {
  pwr_ <- pwr.t.test(d           = i,
                     sig.level   = 0.05,
                     power       = 0.8,
                     type        = "one.sample",
                     alternative = "greater")
  ptab <- rbind(ptab, cbind(pwr_$d, pwr_$n))
}

plot(ptab[,1],
     ptab[,2],
     type="b",
     xlab="effect size",
     ylab="sample size", 
     main="Figure 8: Changes to sample size from incremental effect sizes for comparison analysis",
     cex.main=0.9)

# welch t-test: unequal variance

t <- t.test(female_data$intrinsic_mean, non.female_data$intrinsic_mean)
t.test_results <- tibble("cohen's d" = d,
                         "welch's t" = t$statistic,
                         "p-value"   = t$p.value)

```

Paired-samples Spearman's Rank Correlation.

```{r cor_analysis}

# spearman: ordinal data
s <- cor.test(female_data$intrinsic_mean, female_data$help_mean, method="spearman",
              conf.level=0.95, 
              alternative="two.sided")

spearman_results <- tibble("rho" = s$estimate,
                           "p-value" = s$p.value)

effect_sizes = c(seq(0.1,1.5,0.02))
ptab_s <- cbind(NULL, NULL) 

for (i in effect_sizes) {
  pwr_ <- pwr.t.test(d           = i,
                     sig.level   = 0.05,
                     power       = 0.8,
                     type        = "one.sample",
                     alternative = "two.sided")
  ptab_s <- rbind(ptab_s, cbind(pwr_$d, pwr_$n))
}

plot(ptab_s[,1],
     ptab_s[,2],
     type="b",
     xlab="effect size",
     ylab="sample size", 
     main="Figure 9: Changes to sample size from incremental effect sizes for correlation analysis",     
     cex.main=0.9)

```



